---
title: "p8105_hw5_yh3824"
author: "ruby"
date: "2025-11-11"
output: github_document
---

load packages/set formatting 

```{r, message = FALSE}
library(tidyverse)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

## Problem 1

make a function for birthday simulation 

```{r}
bday_sim = function(n_room) {
  
  birthdays = sample(1:365, n_room, replace = TRUE)


repeated_bday = length(unique(birthdays)) < n_room

repeated_bday
  
}
```

test out the function
```{r}
bday_sim(25)
```

iterate 

```{r}
bday_sim_results = 
  expand_grid(
    bdays = 2:50,
    iter = 1:10000
  ) |> 
  mutate(
    result = map_lgl(bdays, bday_sim)
  ) |> 
  group_by(bdays) |> 
  summarize(
    prob_repeat = mean(result)
  )
```

Make a plot

```{r}
bday_sim_results |> 
  ggplot(aes(x = bdays, y = prob_repeat)) +
  geom_point() +
  geom_line() +
  labs(
    x = "Group Size", 
    y = "Probability of Repeat",
    title = "Probability of Sharing a Repeated Birthday over Increasing Group Size"
  )
```

### Comments

This plot shows us that as the probability of sharing a repeated birthday increases as the sample group increases. This is as expected as the more people who are sampled, the more likely there will be at least 2 people who share the same birthday. At around 23-24 is when we see the probability exceeding 50%, which demonstrates the birthday paradox. 

## Problem 2

for mu = 0

```{r}
powermu_zero = 
  expand_grid(
    sample_n = 30,
    sample_sd = 5,
    mu = 0,
    iter = 1:5000
  ) |> 
  mutate(data = map(iter, ~rnorm(sample_n, mean = mu, sd = sample_sd))) |> 
  mutate(tidy = map(data, ~broom::tidy(t.test(.x, mu = 0)))) |> 
  mutate(mu_hat = map_dbl(tidy, "estimate"),
         p_value = map_dbl(tidy, "p.value"))

```

for mu = {1,2,3,4,5,6}

```{r}
power_sim = 
  expand_grid(
    mu = 1:6, 
    iter = 1:5000) |> 
  mutate(
    data = map(mu, ~ rnorm(n = 30, mean = .x, sd = 5)),
    ttest = map(data, ~broom::tidy(t.test(.x, mu = 0)))
    
  ) |> 
  unnest(ttest) |> 
  select(mu, iter, estimate, p.value)
  
```


make a plot for proportion of times the null was rejected (power) and the true value of mu

```{r message = FALSE}
power_sim |> 
  group_by(mu) |> 
  summarize(
    power = mean(p.value < 0.05)
  ) |> 
  ggplot(aes(x = mu, y = power)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
  x = "True Value of μ",
  y = "Power (Proportion Null Rejected)",
  title = "Power and the True Value of μ"
)
  
```

### Comments

This plot shows that the power of the test increases as the true value of μ increases. When μ is close to the null (0), the test rarely rejects the null, resulting in low power. As μ increases, the effect becomes easier to detect, and the probability of rejecting the null increases.



make a plot with avg estimate of mu and true value of mu

```{r}
power_sim |> 
  group_by(mu) |> 
  summarize(
    avg_est = mean(estimate),
    avg_est_rejectnull = mean(estimate[p.value < 0.05], na.rm = TRUE)
  ) |> 
  pivot_longer(cols = c(avg_est, avg_est_rejectnull),
               names_to = "type",
               values_to = "value") |> 
  ggplot(aes(x = mu, y = value, color = type)) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d(labels = c("All Estimates", "Estimates (Null Rejected)")) +
  labs(
    x = "True Value of μ",
    y = "Average Estimate of μ",
    color = "Estimate Type",
    title = "Average Estimates of μ vs. True μ"
  )
```

### Comments

Across the samples, the average estimate of mu_hat is reasonably similar to the true value of mu, reflecting an unbiased estimator. However, when looking only at samples where the null was rejected, the average mu_hat is larger for smaller true mu values. This could be because as mu becomes larger, the probability of the test rejecting the null is higher, so the average estimate would match up closer to the true mu. 


## Problem 3

import homicide-csv data 
```{r message = FALSE} 
homicide_df = read_csv(file = "./data/homicide-data.csv")
```

### Description of the Raw Data 

The `homicide_df` dataset has `r nrow(homicide_df)` observations, with `r ncol(homicide_df)` variables. The dataset describes homicides in 50 large U.S. cities that occurred between approximately 2007-2017. It includes information including where the homicide occurred, the victim name, race, sex, and age. It also includes disposition of the homicide - if it is Open/no arrest, closed by arrest, and closed without arrest. 


Create a `city_state` variable 

```{r}
homicide_df = 
  homicide_df |> 
    mutate(
      city_state = paste(city, state, sep = ", ")
    ) 
```

write a `summarize_homicides` function that summarizes the total homicides, and total unsolved homicides

```{r}
summarize_homicides = function(df, disposition_col = "disposition",
                                unsolved_values = c("Closed without arrest", "Open/No arrest")) {
  
  df |> 
    summarize(
      total_homicide = n(),
      unsolved_homicide = sum(.data[[disposition_col]] %in% unsolved_values)
    )
}
```

For the city of Baltimore, MD, pull the estimate proportions and CI of unsolved homicides

```{r}
baltimore_proptest = 
  homicide_df |> 
  filter(city_state == "Baltimore, MD") |> 
  summarize_homicides() |> 
 (\(df) prop.test(
    x = pull(df, unsolved_homicide),
    n = pull(df, total_homicide)
  ))() |> 
  broom::tidy() |> 
  select(estimate, conf.low, conf.high)

baltimore_proptest |> 
  knitr::kable(digits = 2)
```

Pull proportion of unsolved homicides, and CI across all cities
```{r}
city_proptest = 
  homicide_df |> 
  group_by(city_state) |> 
  summarize_homicides() |> 
  mutate(
    test_result = map2(
      unsolved_homicide,
      total_homicide, ~ prop.test(.x, .y)  |>  
    broom::tidy())) |> 
  unnest(test_result) |> 
  select(city_state, estimate, conf.low, conf.high)

city_proptest |> 
  knitr::kable(digits = 2)
```

create a plot

```{r}
city_proptest |> 
    arrange(estimate) |> 
    mutate(
      city_state = factor(city_state, levels = city_state)
      ) |> 
    ggplot(aes(x = city_state, y = estimate)) +
    geom_point(size = 1) +                      
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),       
                  width = 0.2) +
    labs(
      x = "City",
      y = "Proportion of Unsolved Homicides",
      title = "Estimated Proportion of Unsolved Homicides by City",
      subtitle = "With 95% Confidence Intervals"
    ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

### Comments

Proportion of unsolved homicides is highest in Chicago, IL, and the lowest in Richmond, VA. (Tulsa, AL is the lowest at 0, this might be an error in the dataset)