---
title: "p8105_hw5_yh3824"
author: "ruby"
date: "2025-10-30"
output: github_document
---

load packages/set formatting 

```{r, message = FALSE}
library(tidyverse)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

### Problem 1

```{r}
birthdays = sample(1:365, 5, replace = TRUE)


repeated_bday = length(unique(birthdays)) < 5

repeated_bday
```

put this into a function 

```{r}
bday_sim = function(n_room) {
  
  birthdays = sample(1:365, n_room, replace = TRUE)


repeated_bday = length(unique(birthdays)) < n_room

repeated_bday
  
}
```

test out the function
```{r}
bday_sim(25)
```

iterate 

```{r}
bday_sim_results = 
  expand_grid(
    bdays = 2:50,
    iter = 1:10000
  ) |> 
  mutate(
    result = map_lgl(bdays, bday_sim)
  ) |> 
  group_by(bdays) |> 
  summarize(
    prob_repeat = mean(result)
  )
```

plot this 

```{r}
bday_sim_results |> 
  ggplot(aes(x = bdays, y = prob_repeat)) +
  geom_point() +
  geom_line() +
  labs(
    x = "Group Size", 
    y = "Probability of Repeat",
    title = "Probability of Sharing a Repeated Birthday over Increasing Group Size"
  )
```

This plot shows us that as the probability of sharing a repeated birthday increases as the sample group increases. This is as expected as the more people who are sampled, the more likely there will be at least 2 people who share the same birthday. At around 23-24 is when we see the probability exceeding 50%, which demonstrates the birthday paradox. 

### Problem 2

for mu = 0

```{r}
prob2 = 
  expand_grid(
    sample_n = 30,
    sample_sd = 5,
    mu = 0,
    iter = 1:5000
  ) |> 
  mutate(data = map(iter, ~rnorm(sample_n, mean = mu, sd = sample_sd))) |> 
  mutate(tidy = map(data, ~broom::tidy(t.test(.x, mu = 0)))) |> 
  mutate(mu_hat = map_dbl(tidy, "estimate"),
         p_value = map_dbl(tidy, "p.value"))

  
```

for mu = {1,2,3,4,5,6}
```{r}
power_sim = 
  expand_grid(
    mu = 1:6, 
    iter = 1:5000) |> 
  mutate(
    data = map(mu, ~ rnorm(n = 30, mean = .x, sd = 5)),
    ttest = map(data, ~broom::tidy(t.test(.x, mu = 0)))
    
  ) |> 
  unnest(ttest) |> 
  select(mu, iter, estimate, p.value)
  
```


make a plot for proportion of times the null was rejected (power) and the true value of mu

```{r}
power_sim |> 
  group_by(mu) |> 
  summarize(
    power = mean(p.value < 0.05)
  ) |> 
  ggplot(aes(x = mu, y = power)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
  x = "True Value of μ",
  y = "Power (Proportion Null Rejected)",
  title = "Power and the True Value of μ"
)
  
```
insert write up! 

make a plot with avg estimate of mu and true value of mu

```{r}
power_sim |> 
  group_by(mu) |> 
  summarize(
    avg_est = mean(estimate),
    avg_est_rejectnull = mean(estimate[p.value < 0.05], na.rm = TRUE)
  ) |> 
  pivot_longer(cols = c(avg_est, avg_est_rejectnull),
               names_to = "type",
               values_to = "value") |> 
  ggplot(aes(x = mu, y = value, color = type)) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d(labels = c("All Estimates", "Estimates (Null Rejected)")) +
  labs(
    x = "True Value of μ",
    y = "Average Estimate of μ",
    color = "Estimate Type",
    title = "Average Estimates of μ vs. True μ"
  )
```


Insert write up! 

### Problem 3

import homicide-csv data 
```{r}
homicide_df = read_csv(file = "./data/homicide-data.csv")
```

## Description of the Data 

Create a `city_state` variable 

```{r}
homicide_df = 
  homicide_df |> 
    mutate(
      city_state = paste(city, state, sep = ", ")
    ) 
```

write a `summarize_homicides` function

```{r}
summarize_homicides = function(df, disposition_col = "disposition",
                                unsolved_values = c("Closed without arrest", "Open/No arrest")) {
  
  df |> 
    summarize(
      total_homicide = n(),
      unsolved_homicide = sum(.data[[disposition_col]] %in% unsolved_values)
    )
}
```

For the city of Baltimore, MD 
* use `prop.test` to estimate the proportion of homicides that are unsolved, 
* save `prop.test` output as an R object, 
* apply `broom::tidy` and 
* pull the estimate propotion and CI

```{r}
baltimore_proptest = 
  homicide_df |> 
  filter(city_state == "Baltimore, MD") |> 
  summarize_homicides() |> 
 (\(df) prop.test(
    x = pull(df, unsolved_homicide),
    n = pull(df, total_homicide)
  ))() |> 
  broom::tidy() |> 
  select(estimate, conf.low, conf.high)
```

map across all cities
```{r}
city_proptest = 
  homicide_df |> 
  group_by(city_state) |> 
  summarize_homicides() |> 
  mutate(
    test_result = map2(
      unsolved_homicide,
      total_homicide,
      ~ prop.test(.x, .y)  |>  
    broom::tidy())) |> 
  unnest(test_result) |> 
  select(city_state, estimate, conf.low, conf.high)
```

create a plot

```{r}
city_proptest |> 
    arrange(estimate) |> 
    mutate(
      city_state = factor(city_state, levels = city_state)
      ) |> 
    ggplot(aes(x = city_state, y = estimate)) +
    geom_point(size = 1) +                      
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),       
                  width = 0.2) +
    coord_flip() +                                              
    labs(
      x = "City",
      y = "Proportion of Unsolved Homicides",
      title = "Estimated Proportion of Unsolved Homicides by City",
      subtitle = "With 95% Confidence Intervals"
    ) +
    theme_minimal()
```

